Molecular dynamics simulation of synchronization in driven particles

### 2020-11-05 manuscript/manuscript_ajp_2020.tex
updated introduction text

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  Using 
  numerical simulations
  we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  The molecular dynamics simulations
  model the particle dynamics 
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We observe
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  kink propagation on these surfaces,
  and the limits of applied force that
  cause the particles to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

%-----------------------------------------------
%environments
Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as wires and fluid microchannels.
Biological systems such as
neuron axons and capillaries can also be studied
with these models [more detail and references].
Many such systems execute local oscillations
about stable points [elaborate].
%----------------------------------------------------------
%external forces
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

%-------------------------------------------------------------
%experimental - colloids specifically
Colloidal particles trapped in light fields have proven
a particularly useful medium for studying these behaviors.%cite 
The relatively large size of the colloids
and ease in control has lead to a rich array of 
experimental results. %THIS IS A TERRIBLE SENTENCE.
Such studies are considered model systems
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

%------------------------------------------------------------------
%synchronization
Synchronization has been studied for over three centuries,
and is observed in different forms
from Huygens pendulum clocks
to the rhythmic beats of the flapping wings in a flock of birds.
Controlling
synchronization phenomena in weakly coupled oscillators
can be achieved with an external driving force
that causes the syncing of natural oscillation frequencies,
dynamic phase locking \cite{juniper2015}.
[useful and interesting because?]

@@

Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
\vec{F}_{D}(t) = [F_{DC} + F_{AC} \sin(\omega t)] \hat{y},
\end{equation}

with modifiable parameters including
a constant component $F_{DC}$,
and a time dependent component with amplitude $F_{AC}$
and frequency $\omega = 2 \pi f$.
The frequency is scaled in time units,
where [finish].
 We demonstrate how a single particle (Sec.~\ref{sec:one})
 and many particles
 move (Sec.~\ref{sec:twenty})
 in response to this applied force in a variety of light fields.
 In Sec.~\ref{sec:kink} we set $F_{DC}$
 to zero and track the motion
 of a high density area of a particle chain
 (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.

>>>>

  %What problem did you study and why is it important?
  We study synchronization of interacting particles
  confined to a narrow channel 
  driven by an externally applied force.  
  %What methods did you use?
  Using 
  numerical simulations
  we control the particle interactions,
  external force and particle environment
  in order to mimick experimental studies of driven colloidal particles
  confined by light-fields.
  The molecular dynamics simulations
  model the particle dynamics 
  using overdamped equations of motion suitable for a viscous suspension
  of microscopic particles.
  %What were your principal results?
  We observe
  particle synchronization under a variety of conditions,
  including static and dynamic patterns formed on
  periodic and quasiperiod substrates,
  high density kink propagation.
  We note 
  the limits of applied force that
  cause the dynamics to transition to chaotic behavior.
  %What conclusions have you drawn from these results?
  We demonstrate the transition from
  trapped to sliding dynamics in
  quasiperiodic landscapes differs
  from that of a periodic landscape by [TBD],
  and that kinks of high density propagate
  [more/less ?  ] when [what?].
  We include sample code
  and exercises for students
  that include 
  opportunties
  to reproduce our results and propose
  new numerical experiments.
  With only a few particles in two-dimensions,
  the simulation runs quickly,
  making this an appropriate model for undergraduates to explore.

@@

%------------------------------------------------------------------
%synchronization
Synchronization is a universal phenomena. 
Christiaan Huygens is credited with its discovery
with his 1665 observation of
in-phase oscillation of pendulum clocks \cite{Pikovsky2003}.  
Huygens explanation
that the clocks vibrations are coupled through the supporting
wall has since been rigorously demonstrated by 20th century
researchers \cite{Ramirez2016}.
Examples of synchronization occur
in many naturally oscillating systems,
from fireflies syncing with a pulsed LED \cite{},
to simultaneous flapping of birds wings \cite{},
to neuronal firing cardiac muscle \cite{} and brain tissue \cite{}.

%phase locking
Dynamical mode locking is a controlled 
synchronization phenomena in 
which an external driving force
causes the syncing of natural oscillation frequencies
in
weakly coupled oscillators.
Mode locking is often observed in quantum electronic
devices such as Josephson junctions \cite{}
and phase-locked loops \cite{}.
Instead of ohmic current-voltage (IV) relationships,
such devices exhibit stepped regions
such as Shapiro steps \cite{} [explain].

Recently studies of colloidal particles confined
in optical traps have
been used to examine the microscopic dynamics
of mode-locking when subject to external driving force \cite{juniper2015}.

%-----------------------------------------------
%environments
Particles in confined geometries behave differently than free particles.
Stabilized charged particles form patterns
due to the interplay of the confining environment
and particle interactions.
Narrow channels studies are useful to provide insights 
of how particles move through systems 
such as wires and microchannels.
Biological systems such as
neuron axons and capillaries can also be studied
with these models [more detail and references].
Many such systems execute local oscillations
about stable points [elaborate].
%----------------------------------------------------------
%external forces
An applied external force
increases the diversity of behaviors,
and can cause particles to flow in
a variety of non-linear complex behaviors
including
synchronized, aperiod, or chaotic dynamical patterns.
%
The presence of a modulating surface
can modify these patterns in a variety of ways,
changing the onset of dynamical flows,
and the overall flow patterns.

%-------------------------------------------------------------
%experimental - colloids specifically
Colloidal particles trapped in light fields have proven
a particularly useful medium for studying these behaviors.  %cite 
The relatively large size of the colloids
and ease in control has lead to a rich array of 
experimental results. %THIS IS A TERRIBLE SENTENCE.
Such studies are considered model systems
for experimental systems relatively hard to access and visualize,
such as cold atoms or electron gases. %ELABORATE AND CITE

@@

Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
\vec{F}_{D}(t) = [F_{DC} + F_{AC} \sin(\omega t)] \hat{y},
\end{equation}
with modifiable parameters including
a constant component $F_{DC}$,
and a time dependent component with amplitude $F_{AC}$
and frequency $\omega = 2 \pi f$.
The frequency is scaled in time units,
where [finish].
 We demonstrate how a single particle (Sec.~\ref{sec:one})
 and many particles
 move (Sec.~\ref{sec:twenty})
 in response to this applied force in a variety of light fields.
 In Sec.~\ref{sec:kink} we set $F_{DC}$
 to zero and track the motion
 of a high density area of a particle chain
 (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
